---
title: "Assignment 1"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1 (5 points)

For each of parts (a) through (d), indicate whether we would generally expect the
performance of a flexible statistical learning method to be better or worse than an inflexible
method. Justify your answer.

1. The sample size $n$ is extremely large, and the number of predictors $p$ is small.
1. The number of predictors $p$ is extremely large, and the number of observations $n$ is small.
1. The relationship between the predictors and response is highly non-linear.
1. The variance of the error terms, is extremely high.

## Exercise 2 (5 points)

Describe the difference between a parametric and non-parametric statistical learning approach. What are the advantages of a parametric approach to regression or classification (as opposed to a noon-parametric approach)? What are its disadvantages?

## Exercise 3 (5 points)

Carefully explain the the difference between the KNN classifier and KNN regression methods. Name a downside when using this model on very large data.

## Exercise 4 (5 points)

Suppose we have a data set with five predictors, $X1 =$ GPA, $X2 =$ extracurricular activities (EA), $X3 =$ Gender (1 for Female and 0 for Male), $X4 =$ Interaction between GPA and EA, and $X5 =$ Interaction between GPA and Gender. The response is starting salary after graduation (in thousands of dollars). Suppose we use least squares to fit the model, and get $\beta_0 = 50$, $\beta_1 = 20$, $\beta_2 = 0.07$, $\beta_3 = 35$, $\beta_4 = 0.01$, $\beta_5 = - 10$.

a. Which answer is correct, and why?
    1. For a fixed value of EA and GPA, males earn more on average than females.
    1. For a fixed value of EA and GPA, females earn more on average than males.
    1. For a fixed value of EA and GPA, males earn more on average than females provided that the GPA is high enough.
    1. For a fixed value of EA and GPA, females earn more on average than males provided that the GPA is high enough.
a. Predict the salary of a female with EA of 110 and a GPA of 4.0.
b. True or false: Since the coefficient for the GPA/EA interaction term is very small, there is very
little evidence of an interaction effect. Justify your answer.

## Exercise 5 (10 points)

This question should be answered using the `biomass` data set.

```{r, eval=FALSE}
library(tidymodels)
data("biomass")
```

a. Fit a multiple regression model to predict `HHV` using `carbon`, `hydrogen` and `oxygen`.
a. Provide an interpretation of each coefficient in the model. Be careful, note the values `Cruise` is able to take.
a. Write out the model in equation form.
a. For which the predictors can you reject the null hypothesis $H_0: \beta_j = 0$?
a. On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.
a. How well do the models in (a) and (e) fit the data? How big was the effect of removing the predictor?
