---
title: "Assignment 6"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1

Explain the assumptions we are making when performing Principle Component Analysis (PCA). What happens when these assumptions are violated?

# Exercise 2

Answer the following questions regarding Principle Component Analysis.

- Is it important to standardize before applying PCA?
- Should one remove highly correlated variables before doing PCA?
- What will happen when eigenvalues are roughly equal?
- Can PCA be used to reduce the dimensionality of a highly nonlinear data set?

# Exercise 3

You will in this exercise explore a data set using PCA. The data comes from the [#tidytuesday project](https://github.com/rfordatascience/tidytuesday) and is about [Student Loan Payments](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-11-26).

Load in the data using the following script.

```{r, eval=FALSE}
library(tidymodels)
loans <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-26/loans.csv") %>%
  select(-agency_name, -added) %>%
  drop_na()
```

a.  Use the `prcomp()` function to perform PCA on the loans data set. Set `scale. = TRUE` to perform scaling. What results are contained in this object? (hint: use the `names()` function)

b. Calculate the amount of variance explained by each principal component. (hint: look at `?broom:::tidy.prcomp`)

c. Use the `tidy()` function to extract the the loadings. Which variable contributed most to the first principle component? Second Component?

d. Use the `augment()` function to get back the transformation and create a scatter plot of any two components of your choice.

# Exercise 4

In this exercise, you are tasked to predict the weight of an animal in a zoo, based on which words are used to describe it. The `animals` data set can be downloaded [here](data/animals.csv).

This data set contains 1001 variables. The first variable `weight` is the natural log of the mean weight of the animal. The remaining variables are named `tf_*` which shows how many times the word `*` appears in the description of the animal.

Use {tidymodels} to set up a workflow to train a PC regression. We can do this by specifying a linear regression model, and create a preprocessor recipe with {recipes} that applies PCA transformation on the predictors using `step_pca()`. Use the `threshold` argument in `step_pca()` to only keep the principal components that explain 90% of the variance.

How well does this model perform on the testing data set?
