---
title: "Assignment 2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1 (10 points)

Suppose we collect data for a group of students in a statistics class with variables $X_1$ = hours studied, $X_2$ = undergrad GPA, and $Y$ = receive an A. We fit a logistic regression and produce estimated coefficient, $\hat{\beta}_0=-6$, $\hat{\beta}_1=0.05$, $\hat{\beta}_2=1$.

a) Estimate the probability that a student who studies for 40 hours and has an undergrad GPA of $3.5$ gets an A in the class.
a) How many hours would that student in part (a) need to study to have a 50% chance of getting an A in the class?

# Exercise 2 (5 points)

Suppose that we take a data set, divide it into equally-sized training and test sets, and then try out two different classification procedures. First, we use logistic regression and get an error rate of 20% on the training data and 30% on the test data. Next, we use 1-nearest neighbors (i.e. $K = 1$) and get an average error rate (averaged over both test and training data sets) of 18%. Based on these results, which method should we prefer to use for classification of new observations? Why?

# Exercise 3 (15 points)

In this exercise, we will explore a data set about cars called `auto` which you can find [here](data/auto.csv).

The data set contains 1 factor variable and 6 numeric variables. The factor variable `mpg` has two levels `high` and `low` indicating whether the car has a high or low miles per gallon. We will in this exercise investigate if we are able to use a logistic regression classifier to predict if a car has high or low mpg from the other variables.

a) Read in the data and create a test-train `rsplit` object of `auto` using `initial_split()`. Use default arguments for `initial_split()`.

b) Create the training and testing data set with `training()` and `testing()` respectively.

c) Fit a logistic regression model using `logistic_reg()`. Use all the 6 numeric variables as predictors (a formula shorthand is to write `mpg ~ .` where `.` means everything. Remember to fit the model only using the training data set.

d) Inspect the model with `summary()` and `tidy()`. Which of the variables are significant?

e) Predict values for the training data set and save them as `training_pred`.

f) Use the following code to calculate the training accuracy

```{r, eval=FALSE}
bind_cols(
  training_pred,
  auto_training
) %>%
  accuracy(truth = mpg, estimate = .pred_class)
```

(`auto_training` should be renamed to match your training data set if needed.)

g) Predict values for the testing data set and use the above code to calculate the testing accuracy. Compare.
