---
title: "Week 5 - Feature Engineering"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Please read [Feature engineering with recipes](https://www.tmwr.org/recipes.html) from Tidy Modeling with R.

[recipes](https://recipes.tidymodels.org/) will be introduces as the R package we will use to perform pre-processing. 

As optional but highly recommended reading is chapter 5, 6, and 8 from [Feature Engineering and Selection: A Practical Approach for Predictive Models](http://www.feat.engineering/).

- [Encoding Categorical Predictors](http://www.feat.engineering/encoding-categorical-predictors.html)
- [Engineering Numeric Predictors](http://www.feat.engineering/engineering-numeric-predictors.html)
- [Handling Missing Data](http://www.feat.engineering/handling-missing-data.html)

## Slides

```{r, echo=FALSE, eval=TRUE}
xaringanExtra::embed_xaringan(url = "static/slides/week-05/index.html", ratio = "16:9")
```

[PDF download](static/slides/week-05/week-05.pdf) [Full Screen](static/slides/week-05/index.html)
