---
title: "Week 3 - Logistic Regression"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Start read chapter 4 of "An Introduction to Statistical Learning", we will color 4.1, 4.2, and 4.3. This chapter covers logistic regression, linear discriminant analysis, quadratic discriminant analysis, and Naive Bayes.
We will start by talking about logistic regression this week and talk about linear discriminant analysis and quadratic discriminant analysis next week.
Furthermore, this is the first we will be talking about classification methods, take note of the differences between this and regression.

We will also talk about the notion of data splitting or "train-test splitting". Please read the 5th chapter of "Tidy Modeling with R" which can be found [here](https://www.tmwr.org/splitting.html).

The reading labs for chapter 4 can be found [here](https://emilhvitfeldt.github.io/ISLR-tidymodels-labs/classification.html)

## Slides

```{r, echo=FALSE, eval=TRUE}
xaringanExtra::embed_xaringan(url = "static/slides/week-03/index.html", ratio = "16:9")
```

[PDF download](static/slides/week-03/week-03.pdf) [Full Screen](static/slides/week-03/index.html)
