---
title: "Schedule"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(glue)
library(fs)
library(stringr)

link_generator <- function(name) {
  function(x) {
    if (file_exists(glue("{name}-{x}.Rmd"))) {
      url <- glue("{name}-{x}.html")
      return(glue("[{name}]({url})"))
    } else {
      return(glue("[{str_to_title(name)} haven't been posted yet]()"))
    }
  }
}

readings <- link_generator("readings")
labs <- link_generator("labs")
assignments <- link_generator("assignments")
```

```{r}
library(lubridate)
first_week <- date("2021-08-30") %m+% days(c(0, 6))

week_days <- function(x) {
  days <- first_week %m+% days((x - 1) * 7)
  glue("*{days[1]} to {days[2]}*")
}
```

# Week 1

`r week_days(1)`

### Monday

Topics: Introductions, What is Statistical Machine Learning?, R, RStudio, Tidymodels

`r readings("01")`

### Thursday

Topics: Linear regression, Regression

# Week 2

`r week_days(2)`

Topics: Linear regression, Regression

`r readings("02")`

`r labs("02")`

`r assignments("01")`

# Week 3

`r week_days(3)`

Topics: Logistic Regression, Classification, Train-Test Split

`r readings("03")`

`r labs("03")`

`r assignments("02")`

# Week 4

`r week_days(4)`

Topics: LDA, QDA, K-Nearest Neighbors, Naive Bayes

`r readings("04")`

`r labs("04")`

`r assignments("03")`

# Week 5

`r week_days(5)`

Topics: Feature Engineering, Data Preprocessing

`r readings("05")`

`r labs("05")`

No assignment due to "All American Weekend"

# Week 6

`r week_days(6)`

Topics: Bootstrap, Model Diagnostics, Evaluation Metrics, Cross Validation

`r readings("06")`

`r labs("06")`

`r assignments("04")`

# Week 7

`r week_days(7)`

Topics: Clustering, K-Means Clustering, Hierarchical Clustering

`r readings("07")`

`r labs("07")`

`r assignments("05")`

# Week 8

`r week_days(8)`

Topics: Topics in ML

`r readings("08")`

`r labs("08")`

`r assignments("midterm")`

# Week 9

`r week_days(9)`

Topics: PCA, PCA Regression

`r readings("09")`

`r labs("09")`

`r assignments("06")`

# Week 10

`r week_days(10)`

Topics: Shrinkage Methods, Rigde, Lasso, Hyper Parameter Tuning

`r readings("10")`

`r labs("10")`

`r assignments("07")`

# Week 11

`r week_days(11)`

Topics: Splines, GAM

`r readings("11")`

`r labs("11")`

`r assignments("08")`

# Week 12

`r week_days(12)`

Topics: Decision Trees, Bagging, Boosting, Random Forests

`r readings("12")`

`r labs("12")`

`r assignments("09")`

# Week 13

`r week_days(13)`

Topics: SVM

`r readings("13")`

No lab due to Thanksgiving

`r assignments("10")`

# Week 14

`r week_days(14)`

Topics: SVM Labs

`r readings("14")`

`r labs("14")`

`r assignments("final")`

# Week 15

`r week_days(15)`

Topics: Final Project Presentations

`r readings("15")`

No labs this week due to Final

`r assignments("final")`
